timesliced_random_cell_selection

### 2018-11-15 src/timesliced_random_cell_selection.Rmd
convex hull done - new data being added

```{r pre-install & load packages, include=FALSE}
packages <- c("raster", "dplyr", "data.table", "sp", "adehabitatHR")
source("../src/ipak.R")
ipak(packages)
```

@@

```{r}
plot(glo_unique_cell)
```

@@

6729...

ok now you want to generate a centroid for each cell... (a point in the center of each cell)
use the raster layer to 1) transfer the raster into a dataframe, with the values (cell numbers) linked to each centroid with xy coordinates (note as the raster is an aea projection, the xy will be in meters)
```{r}
library(raster)
unique_cell_centroid <- as.data.frame(glo_unique_cell, row.names = NULL, optional = FALSE, xy = TRUE, centroids = TRUE, na.rm = TRUE)
names(unique_cell_centroid)[names(unique_cell_centroid)=="layer.1"] <- "cell_id" #just renames the output
names(unique_cell_centroid)[names(unique_cell_centroid)=="x"] <- "longitude" #just renames the output
names(unique_cell_centroid)[names(unique_cell_centroid)=="y"] <- "latitude" #just renames the output
write.csv(unique_cell_centroid, "../output/env/unique_cell_centroid.csv")
plot(glo_unique_cell)
points(unique_cell_centroid$longitude, unique_cell_centroid$latitude, pch = 10, col = "black")
```
well... the plot doesn't tell me much... but i definately have points!

ok so now I have a base list to work off... This gives you a dataset of xy points, but you need xyz (where z is based on the depth layers in the environmental data)
```{r}
unique_cell_xyz <- unique_cell_centroid
unique_cell_xyz$depthlayerno <- NA
unique_cell_xyz <- unique_cell_xyz[rep(seq_len(nrow(unique_cell_xyz)), each=50),] #this generates 50 repetitions of each row
unique_cell_xyz$depthlayerno <- rep(1:50) #assign 1:50 on a repeating cycle to the depth column. 50 is chosen because there are 50 depth layers
unique_cell_xyz$id_depth <- paste(unique_cell_xyz$cell_id, unique_cell_xyz$depthlayerno, sep="") #create a new column that is a unique cell_id & depth ID. This is to help remove cells later and randomise
```

Split the data_aea into timeslices (year/month).
All you need is the year, month, cell_id, and depthlayerno
```{r}
cell_yymm <- subset(data_aea, select= c(cell_id, depthlayerno, year, month)) #create a new df with just cell_id, yr, month
cell_yymm$id_depth <- paste(cell_yymm$cell_id, cell_yymm$depthlayerno, sep="") #create a new column that is a unique cell_id & depth ID. This is to help remove cells later and randomise
split_cell_yymm <- split(cell_yymm, list(cell_yymm$year, cell_yymm$month), drop = TRUE)
```

@@

ok yea, i guess.... still wondering about those edge areas....anyway

@@

@@

```{r}
library(dplyr)
bckrnd_folder <- "../output/bio/background/raw/"      # path to folder that holds multiple .csv files
bckrnd_list <- list.files(path = bckrnd_folder, pattern="*.csv") # create list of all .csv files in folder
randbckoutput <- "../output/bio/background/rand/" #where the files are to be saved to
# read in each .csv file in file_list and create a data frame with the same name as the .csv file
for (i in 1:length(bckrnd_list)){
  sel <- read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
  sel <- sample_n(test_sel, 10000) #where 10000 = number of rows to sample (large sample as per maxent)
  write.csv(sel, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
}

```

@@

```{r}
rand_folder <- "../output/bio/background/rand/"      # path to folder that holds multiple .csv files
rand_list <- list.files(path = rand_folder, pattern="*.csv") # create list of all .csv files in folder
randbckoutput <- "../output/bio/background/rand/" #where the files are to be saved to
for(i in 1:length(rand_list)){
  randtemp <- read.csv(paste(rand_folder, rand_list[i], sep=''))
  randtemp$year <- substr(rand_list[i], 1, 4) #the year is characters 1-4 in the filename
  randtemp$month <- substr(rand_list[i], 6, 7) #the month is characters 6 & 7 in the filename (except where the month is a single digit)
  write.csv(randtemp, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
}
```
note that in the output temp file in R shows a . after the month if that month is a single digit. The .csv outputs don't

@@

```{r}
library(data.table)
filenames  <- list.files(path = rand_folder, pattern="*.csv", full.names = TRUE) 
background_all <- rbindlist(lapply(filenames,fread))
write.csv(background_all, "../output/bio/background_all.csv", row.names = FALSE)
```

@@

Easy first one - add an id based on id_depth
```{r}
background_all$id <- background_all$id_depth
write.csv(background_all, "../output/bio/background_all.csv", row.names = FALSE)
```

@@

```{r}
library(sp)
background_all <- as.data.frame(background_all)
xy <- background_all[ ,c(1,2)] #lon lat in col 1 and 2
sp_background <- SpatialPointsDataFrame(coords = xy, data = background_all, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"))
head(sp_background)
```

@@

```{r}
sp_background <- data.frame(test, long=coordinates(sp_background)[,1], lat=coordinates(sp_background)[,2])
head(sp_background)
```

@@

```{r NAFO shapefile extraction}
library(rgdal)
library(sp)
coordinates(sp_background) <- c("decimalLongitude", "decimalLatitude") 
nafo_zones <- readOGR(dsn=path.expand("../data/bio/nafo_zones"), layer = "nafo_zones_wgs84") #this loads the shapefile
proj4string(sp_background) <- proj4string(nafo_zones) #tells R that the occurrence data is the same projection as the shapefile
sp_background$nafo_zone <- over(sp_background, nafo_zones)$ZONE #ZONE is where the zone data is held in teh shapefiles' attributes
head(sp_background) #just to check it worked
```

>>>>

New package attempt to do this....
old code 
packages <- c("raster", "dplyr", "data.table", "sp", "adehabitatHR", "maptools")
source("../src/ipak.R")
ipak(packages)

```{r pre-install & load packages, include=FALSE}
install.packages("pacman")
library(pacman)
p_load(raster, dplyr, data.table, sp, adehabitatHR, rgdal)
```

@@

```{r}
glo_unique_cell <- raster("../output/env/glo_unique_cell.tif") #this is loading the cell layer that was created and saved as a tif
plot(glo_unique_cell)
```

@@

7750

ok now you want to generate a centroid for each cell... (a point in the center of each cell)
use the raster layer to 1) transfer the raster into a dataframe, with the values (cell numbers) linked to each centroid with xy coordinates (note as the raster is an aea projection, the xy will be in meters)
```{r}
unique_cell_centroid <- as.data.frame(glo_unique_cell, row.names = NULL, optional = FALSE, xy = TRUE, centroids = TRUE, na.rm = TRUE)
names(unique_cell_centroid)[names(unique_cell_centroid)=="layer.1"] <- "cell_id" #just renames the output
names(unique_cell_centroid)[names(unique_cell_centroid)=="x"] <- "longitude" #just renames the output
names(unique_cell_centroid)[names(unique_cell_centroid)=="y"] <- "latitude" #just renames the output
write.csv(unique_cell_centroid, "../output/env/unique_cell_centroid.csv")
plot(glo_unique_cell)
points(unique_cell_centroid$longitude, unique_cell_centroid$latitude, pch = 10, col = "black")
write.csv(unique_cell_centroid, "../output/env/unique_cell_points.csv", row.names = FALSE)
```
well... the plot doesn't tell me much... but i definately have points!

@@

ok yea, i guess.... still wondering about those edge areas....anyway save the hull as a polygon

```{r}
writeOGR(mcp_capelin_100, dsn = '.', layer = '../output/bio/mcp_capelin_100_poly', driver = "ESRI Shapefile")
```

now need to select points within a polygon....
```{r}
unique_cell_centroid <- read.csv("../output/env/unique_cell_centroid.csv",header = TRUE)
unique_cell_xy <- unique_cell_centroid[ , c("longitude", "latitude")] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
unique_c_sp <- SpatialPointsDataFrame(coords = unique_cell_xy, data = unique_cell_centroid, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
plot(mcp_capelin_100)
points(unique_c_sp, pch = "red")
```

```{r}
mcp_unique_cell <- unique_c_sp[mcp_capelin_100, ] #this is just subsetting 
plot(mcp_capelin_100)
points(mcp_unique_cell)
```
ok now convert mcp_unique_cell back to a dataframe for further processing, and save the clipped unique cells to a .csv
```{r}
mcp_unique_cell <- as.data.frame(mcp_unique_cell)
write.csv(mcp_unique_cell, "../output/env/unique_cell_centroid_mcp")
```

ok so now I have a base list to work off... This gives you a dataset of xy points, but you need xyz (where z is based on the depth layers in the environmental data)
```{r}
unique_cell_xyz <- mcp_unique_cell
unique_cell_xyz$depthlayerno <- NA
unique_cell_xyz <- unique_cell_xyz[rep(seq_len(nrow(unique_cell_xyz)), each=50),] #this generates 50 repetitions of each row
unique_cell_xyz$depthlayerno <- rep(1:50) #assign 1:50 on a repeating cycle to the depth column. 50 is chosen because there are 50 depth layers
unique_cell_xyz$id_depth <- paste(unique_cell_xyz$glo_unique_cell, unique_cell_xyz$depthlayerno, sep="") #create a new column that is a unique glo_unique_cell & depth ID. This is to help remove cells later and randomise
```

@@

Split the data_aea into timeslices (year/month).
All you need is the year, month, cell_id, and depthlayerno
```{r}
data_aea <- mcpbase
cell_yymm <- subset(data_aea, select= c(cell_id, depthlayerno, year, month)) #create a new df with just cell_id, yr, month
cell_yymm$id_depth <- paste(cell_yymm$cell_id, cell_yymm$depthlayerno, sep="") #create a new column that is a unique cell_id & depth ID. This is to help remove cells later and randomise
split_cell_yymm <- split(cell_yymm, list(cell_yymm$year, cell_yymm$month), drop = TRUE)
```

@@

```{r}
bckrnd_folder <- "../output/bio/background/raw/"      # path to folder that holds multiple .csv files
bckrnd_list <- list.files(path = bckrnd_folder, pattern="*.csv") # create list of all .csv files in folder
randbckoutput <- "../output/bio/background/rand/" #where the files are to be saved to
# read in each .csv file in file_list and create a data frame with the same name as the .csv file
for (i in 1:length(bckrnd_list)){
  sel <- read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
  sel <- sample_n(sel, 10000) #where 10000 = number of rows to sample (large sample as per maxent)
  write.csv(sel, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
}
head(sel)  
```

@@

```{r}
rand_folder <- "../output/bio/background/rand/"      # path to folder that holds multiple .csv files
rand_list <- list.files(path = rand_folder, pattern="*.csv") # create list of all .csv files in folder
randbckoutput <- "../output/bio/background/rand/" #where the files are to be saved to
for(i in 1:length(rand_list)){
  randtemp <- read.csv(paste(rand_folder, rand_list[i], sep=''))
  randtemp$year <- substr(rand_list[i], 1, 4) #the year is characters 1-4 in the filename
  randtemp$month <- substr(rand_list[i], 6, 7) #the month is characters 6 & 7 in the filename (except where the month is a single digit)
  write.csv(randtemp, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
}
head(randtemp)
```
note that in the output temp file in R shows a . after the month if that month is a single digit. The .csv outputs don't

@@

```{r}
filenames  <- list.files(path = rand_folder, pattern="*.csv", full.names = TRUE) 
background_all <- rbindlist(lapply(filenames,fread, fill = TRUE))
write.csv(background_all, "../output/bio/background_all.csv", row.names = FALSE)
head(background_all)
```

@@

Easy first one - add an id based on id_depth
```{r}
background_all$id <- background_all$id_depth
write.csv(background_all, "../output/bio/background_all.csv", row.names = FALSE)
head(background_all)
```

@@

```{r}
background_all <- as.data.frame(background_all)
xy <- background_all[ ,c(2,3)] #lon lat in col 1 and 2
sp_background <- SpatialPointsDataFrame(coords = xy, data = background_all, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"))
head(sp_background)
```

@@

```{r}
sp_background <- data.frame(sp_background, long=coordinates(sp_background)[,1], lat=coordinates(sp_background)[,2])
head(sp_background)
```

@@

```{r NAFO shapefile extraction}
coordinates(sp_background) <- c("decimalLongitude", "decimalLatitude") 
nafo_zones <- readOGR(dsn=path.expand("../data/bio/nafo_zones"), layer = "nafo_zones_wgs84") #this loads the shapefile
proj4string(sp_background) <- proj4string(nafo_zones) #tells R that the occurrence data is the same projection as the shapefile
sp_background$nafo_zone <- over(sp_background, nafo_zones)$ZONE #ZONE is where the zone data is held in teh shapefiles' attributes
head(sp_background) #just to check it worked
```
